Eco-evolutionary dynamics of finite populations from first principles

### 2023-02-21 ind_files/06_unification_discussion.tex
Very minor phrasing change in section 6.2

Once again, terms of equation \eqref{stochastic_Price_variance} lend themselves to straightforward biological interpretation. The quantity $(f-\overline{f})^2$ is a measure of the `spread' of the quantity $f$ around the population mean value $\overline{f}$, and thus covariance with $(f-\overline{f})^2$ is a measure of disruptive (\emph{i.e.} diversifying) selection: If values of $f$ that are further away from the mean value are associated with higher fitness (resp. lower $\tau$), the variance of that quantity in the population increases. An extreme case is evolutionary branching, where the mean value $\overline{f}$ is at a local minima for fitness but $f_i \not\equiv \overline{f}$ and thus $\textrm{Cov}(w,(f - \overline{f})^2)$ is strongly positive. In this case, the population experiences a sudden explosion in variance, and, if $\textrm{Cov}(w,(f - \overline{f})^2)$ is still positive after the initial emergence of multiple morphs, evolution eventually leads to the emergence of stable coexisting polymorphisms in the population. The $\textrm{Cov}\left(\partial f/\partial t,f\right)$ term once again represents the effect of eco-evolutionary feedback loops due to rapid change in $f$ that is not solely due to changes in $\mathbf{p}$. The $M_{\sigma^2_f}(\mathbf{p},N_K)$ term quantifies the effect of mutations on the variance of $f$. Note that each $Q_i(\mathbf{p}) \geq 0$ by its definition in \eqref{nD_functional_forms_for_replicator}, and the total population size $KN_K \geq 2$ for all interesting evolutionary questions. Thus, from \eqref{variance_price_mutation_term}, it is clear that $M_{\sigma^2_f}(\mathbf{p},N_K) \geq 0$, meaning that mutations never decrease the variance of $f$ in the population.\\

>>>>

Once again, terms of equation \eqref{stochastic_Price_variance} lend themselves to straightforward biological interpretation. The quantity $(f_i-\overline{f})^2$ is a measure of the distance of $f_i$ from the population mean value $\overline{f}$, and thus covariance with $(f-\overline{f})^2$ is a measure of disruptive (\emph{i.e.} diversifying) selection: If values of $f$ that are further away from the mean value are associated with higher fitness (resp. lower $\tau$), the variance of that quantity in the population increases. An extreme case is evolutionary branching, where the mean value $\overline{f}$ is at a local minima for fitness but $f_i \not\equiv \overline{f}$ and thus $\textrm{Cov}(w,(f - \overline{f})^2)$ is strongly positive. In this case, the population experiences a sudden explosion in variance, and, if $\textrm{Cov}(w,(f - \overline{f})^2)$ is still positive after the initial emergence of multiple morphs, evolution eventually leads to the emergence of stable coexisting polymorphisms in the population. The $\textrm{Cov}\left(\partial f/\partial t,f\right)$ term once again represents the effect of eco-evolutionary feedback loops due to rapid change in $f$ that is not solely due to changes in $\mathbf{p}$. The $M_{\sigma^2_f}(\mathbf{p},N_K)$ term quantifies the effect of mutations on the variance of $f$. Note that each $Q_i(\mathbf{p}) \geq 0$ by its definition in \eqref{nD_functional_forms_for_replicator}, and the total population size $KN_K \geq 2$ for all interesting evolutionary questions. Thus, from \eqref{variance_price_mutation_term}, it is clear that $M_{\sigma^2_f}(\mathbf{p},N_K) \geq 0$, meaning that mutations never decrease the variance of $f$ in the population.\\
